---
title: "Model Comparison"
author: "Andres A Saenz Guzman"
date: "13/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here I compare the forecast accuracy of each model across the five countries in the sample, the US, the EA, the NL, CL and MX. In order to do so I compute the following four metrics: 

(i) Absolute difference in forecast error (AD). Here I take the absolute value of the forecast errors in two models and substract them in order to obtain values and compare them. 

$AD_t = |F_t^i| - |F_t^{-1}|$,

Where $AD$ is the absolute difference, $F_t^i$ is the forecast error in time period $t$ of model $i$ and $F_t^{-i}$ the forecast error in forecast time period $t$ of model $-i$. Note that, if $AD_t > 0$, then model $-i$ is more accurate than model $i$. Finally, in order to see the overall accuracy, the differences are shown in a histogram, this if values are leaning towards positive values, then model $i$ is worse than $-i$.

(ii) Average error. An arithmetic mean of forecast errors.

(ii) Present bias average (PBA). Here I penalize the weight of future forecast errors in the average. The idea is that central banks may be more interested in the prediction accuracy of short term inflation. I compute the average as shown below:

$PBA^i = \frac{\sum_{t=1}^T (1 - \frac{t - 1}{T}) |F_t^i|}{T}$,

Where $PBA^i$ is the present bias average of model $i$, $t$ is the number of time periods in the future, $T$ is the number of periods in the forecast and $F_t^i$ is the forecast error in $t$ of model $i$.

(iii) Present punishing average (PPA). In this average, I do the opposite of the PBA. I penalize short term errors in order to measure the accuracy of longer term forecasts.

$PPA^i =  \frac{\sum_{t=1}^T (1 - \frac{T - t}{T}) |F_t^i|}{T}$,

Where $PPA^i$ is the present punishing average of model $i$, $t$ is the number of time periods in the future, $T$ is the number of periods in the forecast and $F_t^i$ is the forecast error in $t$ of model $i$.

In order to use PBA and PPA, I create a weights array.

```{r}
Te <- length(error.benchmark.us) #since all forecasts are the same length, it does not matter which one I choose 
t <- 1:Te 

w.PBA <- 1 - (t - 1)/Te

w.PPA <- 1 - (Te - t)/Te
```


##Comparing models in the US
```{r US model comparison}
AD.benchmark.var <- abs(error.benchmark.us) - abs(error.var.us)
AD.benchmark.nnetar <- abs(error.benchmark.us) - abs(error.nnetar.us)
AD.var.nnetar <- abs(error.var.us) - abs(error.nnetar.us)


ggplot(as.data.frame(AD.benchmark.var), aes(AD.benchmark.var)) + geom_histogram()
ggplot(as.data.frame(AD.benchmark.nnetar), aes(AD.benchmark.nnetar)) + geom_histogram()
ggplot(as.data.frame(AD.var.nnetar), aes(AD.var.nnetar)) + geom_histogram()
```

```{r US model comparison}
AVG.benchmark <- mean(abs(error.benchmark.us))
PBA.benchmark <- weighted.mean(abs(error.benchmark.us), w.PBA)
PPA.benchmark <- weighted.mean(abs(error.benchmark.us), w.PPA)

AVG.var <- mean(abs(error.var.us))
PBA.var <- weighted.mean(abs(error.var.us), w.PBA)
PPA.var<- weighted.mean(abs(error.var.us), w.PPA)

AVG.nnetar <- mean(abs(error.nnetar.us))
PBA.nnetar <- weighted.mean(abs(error.nnetar.us), w.PBA)
PPA.nnetar <- weighted.mean(abs(error.nnetar.us), w.PPA)



AVG.benchmark
PBA.benchmark
PPA.benchmark

AVG.var
PBA.var
PPA.var

AVG.nnetar
PBA.nnetar
PPA.nnetar
```


##Comparing models in the EA

```{r EA model comparison}
AD.benchmark.var <- abs(error.benchmark.ea) - abs(error.var.ea)
AD.benchmark.nnetar <- abs(error.benchmark.ea) - abs(error.nnetar.ea)
AD.var.nnetar <- abs(error.var.ea) - abs(error.nnetar.ea)


ggplot(as.data.frame(AD.benchmark.var), aes(AD.benchmark.var)) + geom_histogram()
ggplot(as.data.frame(AD.benchmark.nnetar), aes(AD.benchmark.nnetar)) + geom_histogram()
ggplot(as.data.frame(AD.var.nnetar), aes(AD.var.nnetar)) + geom_histogram()
```


```{r EA model comparison}
AVG.benchmark <- mean(abs(error.benchmark.ea))
PBA.benchmark <- weighted.mean(abs(error.benchmark.ea), w.PBA)
PPA.benchmark <- weighted.mean(abs(error.benchmark.ea), w.PPA)

AVG.var <- mean(abs(error.var.ea))
PBA.var <- weighted.mean(abs(error.var.ea), w.PBA)
PPA.var<- weighted.mean(abs(error.var.ea), w.PPA)

AVG.nnetar <- mean(abs(error.nnetar.ea))
PBA.nnetar <- weighted.mean(abs(error.nnetar.ea), w.PBA)
PPA.nnetar <- weighted.mean(abs(error.nnetar.ea), w.PPA)



AVG.benchmark
PBA.benchmark
PPA.benchmark

AVG.var
PBA.var
PPA.var

AVG.nnetar
PBA.nnetar
PPA.nnetar
```


##Comparing models in the NL

```{r NL model comparison}
AD.benchmark.var <- abs(error.benchmark.nl) - abs(error.var.nl)
AD.benchmark.nnetar <- abs(error.benchmark.nl) - abs(error.nnetar.nl)
AD.var.nnetar <- abs(error.var.nl) - abs(error.nnetar.nl)


ggplot(as.data.frame(AD.benchmark.var), aes(AD.benchmark.var)) + geom_histogram()
ggplot(as.data.frame(AD.benchmark.nnetar), aes(AD.benchmark.nnetar)) + geom_histogram()
ggplot(as.data.frame(AD.var.nnetar), aes(AD.var.nnetar)) + geom_histogram()
```


```{r NL model comparison}
AVG.benchmark <- mean(abs(error.benchmark.nl))
PBA.benchmark <- weighted.mean(abs(error.benchmark.nl), w.PBA)
PPA.benchmark <- weighted.mean(abs(error.benchmark.nl), w.PPA)

AVG.var <- mean(abs(error.var.nl))
PBA.var <- weighted.mean(abs(error.var.nl), w.PBA)
PPA.var<- weighted.mean(abs(error.var.nl), w.PPA)

AVG.nnetar <- mean(abs(error.nnetar.nl))
PBA.nnetar <- weighted.mean(abs(error.nnetar.nl), w.PBA)
PPA.nnetar <- weighted.mean(abs(error.nnetar.nl), w.PPA)



AVG.benchmark
PBA.benchmark
PPA.benchmark

AVG.var
PBA.var
PPA.var

AVG.nnetar
PBA.nnetar
PPA.nnetar
```


##Comparing models in the CL

```{r CL model comparison}
AD.benchmark.var <- abs(error.benchmark.cl) - abs(error.var.cl)
AD.benchmark.nnetar <- abs(error.benchmark.cl) - abs(error.nnetar.cl)
AD.var.nnetar <- abs(error.var.cl) - abs(error.nnetar.cl)


ggplot(as.data.frame(AD.benchmark.var), aes(AD.benchmark.var)) + geom_histogram()
ggplot(as.data.frame(AD.benchmark.nnetar), aes(AD.benchmark.nnetar)) + geom_histogram()
ggplot(as.data.frame(AD.var.nnetar), aes(AD.var.nnetar)) + geom_histogram()
```


```{r CL model comparison}
AVG.benchmark <- mean(abs(error.benchmark.cl))
PBA.benchmark <- weighted.mean(abs(error.benchmark.cl), w.PBA)
PPA.benchmark <- weighted.mean(abs(error.benchmark.cl), w.PPA)

AVG.var <- mean(abs(error.var.cl))
PBA.var <- weighted.mean(abs(error.var.cl), w.PBA)
PPA.var<- weighted.mean(abs(error.var.cl), w.PPA)

AVG.nnetar <- mean(abs(error.nnetar.cl))
PBA.nnetar <- weighted.mean(abs(error.nnetar.cl), w.PBA)
PPA.nnetar <- weighted.mean(abs(error.nnetar.cl), w.PPA)



AVG.benchmark
PBA.benchmark
PPA.benchmark

AVG.var
PBA.var
PPA.var

AVG.nnetar
PBA.nnetar
PPA.nnetar
```


##Comparing models in the MX

```{r MX model comparison}
AD.benchmark.var <- abs(error.benchmark.mx) - abs(error.var.mx)
AD.benchmark.nnetar <- abs(error.benchmark.mx) - abs(error.nnetar.mx)
AD.var.nnetar <- abs(error.var.mx) - abs(error.nnetar.mx)


ggplot(as.data.frame(AD.benchmark.var), aes(AD.benchmark.var)) + geom_histogram()
ggplot(as.data.frame(AD.benchmark.nnetar), aes(AD.benchmark.nnetar)) + geom_histogram()
ggplot(as.data.frame(AD.var.nnetar), aes(AD.var.nnetar)) + geom_histogram()
```


```{r MX model comparison}
AVG.benchmark <- mean(abs(error.benchmark.mx))
PBA.benchmark <- weighted.mean(abs(error.benchmark.mx), w.PBA)
PPA.benchmark <- weighted.mean(abs(error.benchmark.mx), w.PPA)

AVG.var <- mean(abs(error.var.mx))
PBA.var <- weighted.mean(abs(error.var.mx), w.PBA)
PPA.var<- weighted.mean(abs(error.var.mx), w.PPA)

AVG.nnetar <- mean(abs(error.nnetar.mx))
PBA.nnetar <- weighted.mean(abs(error.nnetar.mx), w.PBA)
PPA.nnetar <- weighted.mean(abs(error.nnetar.mx), w.PPA)



AVG.benchmark
PBA.benchmark
PPA.benchmark

AVG.var
PBA.var
PPA.var

AVG.nnetar
PBA.nnetar
PPA.nnetar
```