---
title: "Benchmark"
author: "Andres A Saenz Guzman"
date: "13/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, results='hide'}
library(ggplot2)
library(tidyverse)
library(stargazer)
library(forecast)
```


Here I build the benchmark model for the proyect. The model consist in a naive first-order univariate autoregressive process, i.e., an AR(1). This model takes the following form:

$\pi_{t+1} = \alpha + \beta \pi_t + \epsilon_{t+1}$,

Where $\pi_t$ represent inflation at time period $t$, $\alpha$ is a constant, $\beta < 1$ is the persistence and $\epsilon_{t+1}$ is the estimation error.

##Benchmark for the US

```{r AR1 US}
inf <- train_us[, 2]  #Inflation is stored in the second column.
actual.value <- test_us[, 2]
period.future <- test_us[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)

error.benchmark.us <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.us)
```

## Benchmark for the EA

```{r AR1 EA}
inf <- train_ea[, 2]  #Inflation is stored in the second column.
actual.value <- test_ea[, 2]
period.future <- test_ea[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)

error.benchmark.ea <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.ea)
```

## Benchmark for the NL

```{r AR1 NL}
inf <- train_nl[, 2]  #Inflation is stored in the second column.
actual.value <- test_nl[, 2]
period.future <- test_nl[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)

error.benchmark.nl <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.nl)
```

## Benchmark for CL

```{r AR1 CL}
inf <- train_cl[, 2]  #Inflation is stored in the second column.
actual.value <- test_cl[, 2]
period.future <- test_cl[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)

error.benchmark.cl <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.cl)
```

## Benchmark for MX

```{r AR1 MX}
inf <- train_mx[, 2]  #Inflation is stored in the second column.
actual.value <- test_mx[, 2]
period.future <- test_mx[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)

error.benchmark.mx <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.mx)
```


