---
title: "Benchmark"
author: "Andres A Saenz Guzman"
date: "13/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, results='hide'}
library(ggplot2)
library(tidyverse)
library(stargazer)
library(forecast)
library(tseries)
library(scales)
```


Here I build the benchmark model for the proyect. The model consist in a naive first-order univariate autoregressive process, i.e., an AR(1). This model takes the following form:

$\pi_{t+1} = \alpha + \beta \pi_t + \epsilon_{t+1}$,

Where $\pi_t$ represent inflation at time period $t$, $\alpha$ is a constant, $\beta < 1$ is the persistence and $\epsilon_{t+1}$ is the estimation error.

##Benchmark for the US

```{r AR1 US}
inf <- train_us[, 2]  #Inflation is stored in the second column.
actual.value <- test_us[, 2]
period.future <- test_us[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)


plotdate <- as.Date(dataUS$mdate[1:nrow(train_us)])
benchmark.resid <- as.data.frame(benchmark$residuals)
resid.acf <- acf(benchmark.resid)




resid.plot <- ggplot(data=benchmark.resid, aes(x=plotdate, y=x)) + 
  geom_line() + labs(x = 'Residuals', y = 'Date', title ='US inflation residuals from AR(1) model') + theme_bw()
resid.plot

acf.plot <- autoplot(resid.acf, main = 'US AR(1) residuals autocorrelation function') + theme_bw()
acf.plot

resid.dist.plot <- ggplot(data=benchmark.resid, aes(x)) +
  geom_histogram(aes(y=..density.., fill = ..count..)) + 
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(as.ts(benchmark.resid)), 
                          sd=sd(as.ts(benchmark.resid)))) +
  labs(x = 'Residuals', y = 'Density', title = 'US AR(1) residuals distribution') + theme_bw()
resid.dist.plot

norm.test <- jarque.bera.test(as.ts(benchmark.resid))
norm.test

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(as.ts(inf), series = 'Data', 
         main = 'US inflation AR(1) forecast') +
  autolayer(fitted(fcast.benchmark), series = 'Fitted') +
  autolayer(fcast.benchmark, series = 'Forecast') +
  xlab('Time Period') + ylab('Inflation') + theme_bw()
pred.benchmark.us <- fcast.benchmark$mean
error.benchmark.us <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.us, main = 'US AR(1) forecast errors') + 
  xlab('Time Period') + ylab('Prediction error') + theme_bw()
```

## Benchmark for the EA

```{r AR1 EA}
inf <- train_ea[, 2]  #Inflation is stored in the second column.
actual.value <- test_ea[, 2]
period.future <- test_ea[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

plotdate <- as.Date(dataEA$mdate[1:nrow(train_ea)])
benchmark.resid <- as.data.frame(benchmark$residuals)
resid.acf <- acf(benchmark.resid)




resid.plot <- ggplot(data=benchmark.resid, aes(x=plotdate, y=x)) + 
  geom_line() + labs(x = 'Residuals', y = 'Date', title ='EA inflation residuals from AR(1) model') + theme_bw()
resid.plot

acf.plot <- autoplot(resid.acf, main = 'EA AR(1) residuals autocorrelation function') + theme_bw()
acf.plot

resid.dist.plot <- ggplot(data=benchmark.resid, aes(x)) +
  geom_histogram(aes(y=..density.., fill = ..count..)) + 
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(as.ts(benchmark.resid)), 
                          sd=sd(as.ts(benchmark.resid))))  + 
  labs(x = 'Residuals', y = 'Density', title = 'EA AR(1) residuals distribution') + theme_bw()
resid.dist.plot

norm.test <- jarque.bera.test(as.ts(benchmark.resid))
norm.test

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(as.ts(inf), series = 'Data', 
         main = 'EA inflation AR(1) forecast') +
  autolayer(fitted(fcast.benchmark), series = 'Fitted') +
  autolayer(fcast.benchmark, series = 'Forecast') +
  xlab('Time Period') + ylab('Inflation') + theme_bw()

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)
pred.benchmark.ea <- fcast.benchmark$mean
error.benchmark.ea <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.ea, main = 'EA AR(1) forecast errors') + 
  xlab('Time Period') + ylab('Prediction error') + theme_bw()
```

## Benchmark for the NL

```{r AR1 NL}
inf <- train_nl[, 2]  #Inflation is stored in the second column.
actual.value <- test_nl[, 2]
period.future <- test_nl[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

plotdate <- as.Date(dataNL$mdate[1:nrow(train_nl)])
benchmark.resid <- as.data.frame(benchmark$residuals)
resid.acf <- acf(benchmark.resid)




resid.plot <- ggplot(data=benchmark.resid, aes(x=plotdate, y=x)) + 
  geom_line() + labs(x = 'Residuals', y = 'Date', title ='NL inflation residuals from AR(1) model') + theme_bw()
resid.plot

norm.test <- jarque.bera.test(as.ts(benchmark.resid))
norm.test

acf.plot <- autoplot(resid.acf, main = 'NL AR(1) residuals autocorrelation function') + theme_bw()
acf.plot

resid.dist.plot <- ggplot(data=benchmark.resid, aes(x)) +
  geom_histogram(aes(y=..density.., fill = ..count..)) + 
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(as.ts(benchmark.resid)), 
                          sd=sd(as.ts(benchmark.resid)))) + 
  labs(x = 'Residuals', y = 'Density', title = 'NL AR(1) residuals distribution') + theme_bw()
resid.dist.plot

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(as.ts(inf), series = 'Data', 
         main = 'NL inflation AR(1) forecast') +
  autolayer(fitted(fcast.benchmark), series = 'Fitted') +
  autolayer(fcast.benchmark, series = 'Forecast') +
  xlab('Time Period') + ylab('Inflation') + theme_bw()

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)
pred.benchmark.nl <- fcast.benchmark$mean
error.benchmark.nl <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.nl, main = 'NL AR(1) forecast errors') + 
  xlab('Time Period') + ylab('Prediction error') + theme_bw()
```

## Benchmark for CL

```{r AR1 CL}
inf <- train_cl[, 2]  #Inflation is stored in the second column.
actual.value <- test_cl[, 2]
period.future <- test_cl[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

plotdate <- as.Date(dataCL$mdate[1:nrow(train_cl)])
benchmark.resid <- as.data.frame(benchmark$residuals)
resid.acf <- acf(benchmark.resid)




resid.plot <- ggplot(data=benchmark.resid, aes(x=plotdate, y=x)) + 
  geom_line() + labs(x = 'Residuals', y = 'Date', title ='CL inflation residuals from AR(1) model') + theme_bw()
resid.plot

acf.plot <- autoplot(resid.acf, main = 'CL AR(1) residuals autocorrelation function') + theme_bw()
acf.plot

resid.dist.plot <- ggplot(data=benchmark.resid, aes(x)) +
  geom_histogram(aes(y=..density.., fill = ..count..)) + 
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(as.ts(benchmark.resid)), 
                          sd=sd(as.ts(benchmark.resid)))) + 
  labs(x = 'Residuals', y = 'Density', title = 'CL AR(1) residuals distribution') + theme_bw()
resid.dist.plot

norm.test <- jarque.bera.test(as.ts(benchmark.resid))
norm.test

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(as.ts(inf), series = 'Data', 
         main = 'CL inflation AR(1) forecast') +
  autolayer(fitted(fcast.benchmark), series = 'Fitted') +
  autolayer(fcast.benchmark, series = 'Forecast') +
  xlab('Time Period') + ylab('Inflation') + theme_bw()

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)
pred.benchmark.cl <- fcast.benchmark$mean
error.benchmark.cl <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.cl, main = 'CL AR(1) forecast errors') + 
  xlab('Time Period') + ylab('Prediction error') + theme_bw()
```

## Benchmark for MX

```{r AR1 MX}
inf <- train_mx[, 2]  #Inflation is stored in the second column.
actual.value <- test_mx[, 2]
period.future <- test_mx[, 1]

benchmark <- Arima(inf, order = c(1,0,0))
benchmark

checkresiduals(benchmark)

plotdate <- as.Date(dataMX$mdate[1:nrow(train_mx)])
benchmark.resid <- as.data.frame(benchmark$residuals)
resid.acf <- acf(benchmark.resid)




resid.plot <- ggplot(data=benchmark.resid, aes(x=plotdate, y=x)) + 
  geom_line() + labs(x = 'Residuals', y = 'Date', title ='MX inflation residuals from AR(1) model') + theme_bw()
resid.plot

acf.plot <- autoplot(resid.acf, main = 'MX AR(1) residuals autocorrelation function') + theme_bw()
acf.plot

resid.dist.plot <- ggplot(data=benchmark.resid, aes(x)) +
  geom_histogram(aes(y=..density.., fill = ..count..)) + 
  stat_function(fun=dnorm,
                color="red",
                args=list(mean=mean(as.ts(benchmark.resid)), 
                          sd=sd(as.ts(benchmark.resid)))) + 
  labs(x = 'Residuals', y = 'Density', title = 'MX AR(1) residuals distribution') + theme_bw()
resid.dist.plot

norm.test <- jarque.bera.test(as.ts(benchmark.resid))
norm.test

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(as.ts(inf), series = 'Data', 
         main = 'MX inflation AR(1) forecast') +
  autolayer(fitted(fcast.benchmark), series = 'Fitted') +
  autolayer(fcast.benchmark, series = 'Forecast') +
  xlab('Time Period') + ylab('Inflation') + theme_bw()

fcast.benchmark <- forecast(benchmark, h = 12)
autoplot(fcast.benchmark)
pred.benchmark.mx <- fcast.benchmark$mean
error.benchmark.mx <- fcast.benchmark$mean - actual.value
autoplot(error.benchmark.mx, main = 'MX AR(1) forecast errors') + 
  xlab('Time Period') + ylab('Prediction error') + theme_bw()
```


